library(Seurat)
library(dplyr)
library(ggplot2)
library(ggExtra)
library(RColorBrewer)
library(openxlsx)
library(scales)
library(harmony)
library(zellkonverter)
library(SingleCellExperiment)

inputDir <- "C:/Users/gupta/OneDrive/Desktop/input/"
outputDir <- "C:/Users/gupta/OneDrive/Desktop/output/"
plotDir <- "C:/Users/gupta/OneDrive/Desktop/plots/"

rds_files <- list.files(inputDir, pattern = "_seurat\\.rds$", full.names = TRUE)
seurat_list <- lapply(rds_files, readRDS)
names(seurat_list) <- gsub("_seurat\\.rds$", "", basename(rds_files))

seurat_combined <- merge(seurat_list[[1]], y = seurat_list[-1], add.cell.ids = names(seurat_list))
rm(seurat_list)
gc()
metadata <- read.csv("C:/Users/gupta/OneDrive/Desktop/GSE279086metadata - Sheet1.csv", stringsAsFactors = FALSE)
condition_map <- setNames(metadata$Condition, metadata$Sample)
seurat_combined@meta.data$Condition <- unname(condition_map[as.character(seurat_combined@meta.data$sample)])
print(table(seurat_combined$Condition))
saveRDS(seurat_combined, file = paste0(outputDir, "01_GSE279086_seurat_combined.rds"))

dir.create("D:/scRNA/output", recursive = TRUE)
dir.create("D:/scRNA/plots", recursive = TRUE)

outputDir <- "D:/scRNA/output/"
plotDir <- "D:/scRNA/plots/"

saveRDS(seurat_combined, file = "D:/scRNA/output/01_GSE279086_seurat_combined.rds")

gc()
test <- readRDS("D:/scRNA/output/01_GSE279086_seurat_combined.rds")
dim(test)

rm(test)
rm(seurat_combined)
gc()
seurat_combined <- readRDS("D:/scRNA/output/01_GSE279086_seurat_combined.rds")

```{r qc_functions}
filter_nFeatures <- function(seurat_obj, min_feat, max_feat) {
  return(subset(seurat_obj, subset = nFeature_RNA > min_feat & nFeature_RNA < max_feat))
}
filter_mt <- function(seurat_obj, max_mt) {
  return(subset(seurat_obj, subset = percent.mt < max_mt))
}
filter_rb <- function(seurat_obj, max_rb) {
  return(subset(seurat_obj, subset = percent.rb < max_rb))
}
fil_nCounts <- function(seurat_obj, threshold_mahalanobis){
  ncounts_data <- data.frame(nCount_RNA = seurat_obj@meta.data$nCount_RNA)
  mahalanobis_dist <- mahalanobis(x = ncounts_data, center = mean(ncounts_data$nCount_RNA), cov = var(ncounts_data$nCount_RNA))
  seurat_obj$mahal_dist_nCount <- mahalanobis_dist
  mahal.fil <- qchisq(threshold_mahalanobis, df = 1)
  cells_to_remove <- rownames(seurat_obj@meta.data)[seurat_obj$mahal_dist_nCount > mahal.fil]
  subset(seurat_obj, cells = setdiff(colnames(seurat_obj), cells_to_remove))
}

seurat_filtered <- filter_nFeatures(seurat_combined, 200, 7000)
seurat_filtered <- filter_mt(seurat_filtered, 20)
seurat_filtered <- fil_nCounts(seurat_filtered, 0.95)
seurat_filtered <- filter_rb(seurat_filtered, 20)

print(dim(seurat_filtered))
print(table(seurat_filtered$Condition))

saveRDS(seurat_filtered, file = "D:/scRNA/output/02_GSE279086_seurat_qc_filtered.rds")
```
```{r normalize_cluster}
rm(seurat_combined)
gc()

seurat_processed <- NormalizeData(seurat_filtered, normalization.method = "LogNormalize")
seurat_processed <- FindVariableFeatures(seurat_processed, selection.method = "vst", nfeatures = 2500)
seurat_processed <- ScaleData(seurat_processed)
seurat_processed <- RunPCA(seurat_processed, npcs = 50)
seurat_processed <- RunUMAP(seurat_processed, dims = 1:30)
seurat_processed <- FindNeighbors(seurat_processed, dims = 1:30)
seurat_processed <- FindClusters(seurat_processed, resolution = 0.8)

saveRDS(seurat_processed, file = "D:/scRNA/output/03_GSE279086_seurat_pca_umap.rds")
print(dim(seurat_processed))
```
print(dim(seurat_processed))